```{r}
#load data
library(lmtest)
library(leaps)
library(Stat2Data)
bk.data <- read.csv(file = "~/SC321/Reports/Report 4/birdkeeping.csv")
attach(bk.data)
bk.df <- data.frame(bk.data)
print(bk.df)
```
```{r}
#creating table to see how many have lung cancer and are bird keepers
my.table <- table(bk.df$BK, bk.df$LC)
my.table
```
observations:

What are the odds of having a bird? 67/80
What are the odds of having a bird if you have lung cancer? 33/16
What are the odds of having a bird if you don't have lung cancer? 34/64
What are the odds ratio for having a bird comparing those who have and don't have lung cancer? (33/16)/(34/64)
```{r}
#do the math
yesbird <- 67/80
birdcancer <- 33/16
birdnocancer <- 34/64
oddsRatio <- (33/34)/(16/64)
yesbird
birdcancer
birdnocancer
oddsRatio
```
This means that those who have lung cancer are 3.88 times more likely to have a bird than those who do not have lung cancer.

Linear models
```{r}
bk.df$LC_ind <- as.numeric(bk.df$LC == "LUNGCANCER")
bk.df$BK_ind <- as.numeric(bk.df$BK == "BIRD")
main.glm <- glm(bk.df$LC_ind ~ bk.df$BK_ind + as.factor(bk.df$SEX) + as.factor(bk.df$SS) + bk.df$AG + bk.df$YR + bk.df$CD, family="binomial")
#str(bk.df)
summary(main.glm)
```
```{r}
step(main.glm)
```
```{r}
fitMain.glm <- glm(bk.df$LC_ind ~ bk.df$BK_ind + bk.df$YR, family="binomial")
#str(bk.df)
summary(fitMain.glm)
```
```{r}
lrtest(main.glm, fitMain.glm)
```
write description:

Since this p-value is not less than .05, we will fail to reject the null hypothesis.

This means the model and the fitModel fit the data equally well. Thus, we should use the fitModel because the additional predictor variables in the first model don’t offer a significant improvement in fit.




```{r}
library(rpart)
bk.rt <- rpart(bk.df$LC_ind ~ bk.df$BK_ind + bk.df$YR)
par(xpd = TRUE)
plot(bk.rt)
text(bk.rt, pretty=0)
bk.rt
printcp(bk.rt)
```
```{r}
better.glm <- glm(bk.df$LC_ind ~ bk.df$BK_ind + bk.df$YR + I(bk.df$YR^2), family="binomial")
#str(bk.df)
summary(better.glm)
```
```{r}
step(better.glm)
```


```{r}
#creating table to see how many have lung cancer and are bird keepers with new model (odds ratio)
exp(fitMain.glm$coefficients[-1])
```
This means that those who do have lung cancer are 4.37 times more likely to have a bird than those who do not have lung cancer.

```{r}
simple.glm <- glm(bk.df$LC_ind ~ bk.df$BK_ind, family = "binomial")
summary(simple.glm)
```

```{r}
# Confidence interval for crude OR
exp(confint(simple.glm, level = .95))
# Confidence interval for adjusted OR
exp(confint(fitMain.glm, level = .95))
```







